Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-04-14 2_SSL_EST/article/article_supplementary.tex
edits

\section{Luminous Efficacy Metrics}

@@

\newpage
\section{Manufacturing Cost Model}
\label{sec:costmodel}

\subsection{Structure of the Model}

@@

\subsection{Manufacturing Process Steps by Chip Architecture}

@@

\subsection{Computation of Manufacturing Cost}
\label{subsec:cost_computation}

@@

\subsection{Computation of Yielded Cost}

@@

\newpage
\subsection{Example of Input Data Considered: Sapphire Wafers}

@@

\newpage
\subsection{Preliminary Sensitivity Analysis}

@@

\newpage
\subsection{Comparison with Reported Industry Data and DOE Projections}

@@

\newpage
\section{Computation of Contribution of Individual Variables}
\label{sec:contribution_variables}

@@

\section{Non-Phosphor-Related LED Innovations}

@@

\section{Phosphor-Related LED Innovations}
\label{sec:innovation_phosphor}

@@

\subsection{History of Identified Phosphor-Related LED Innovations}

\subsubsection{YAG and YGAG Phosphors}

@@

\subsubsection{258, SLA and SALON phosphors}

@@

\subsubsection{PFS Phosphor}

@@

\subsubsection{Quantum Dots for Light Down-Conversion}

@@

\subsection{Spectral Data for Identified LED Phosphor Innovations}

@@

\section{List of InterviewEST}

@@

>>>>

\section{Metrics}

@@

\subsection{Sub-Efficiencies}

\subsubsection{Forward Voltage Efficiency $\eta_{V_f}$}

Device forward voltage efficiency\footnote{"Joule Efficiency" $\epsilon_J$ in Tsao et al. \cite{tsao2010solid}} (VfE) describes all electrical losses at the interface between the electrodes and the semiconductor and in the bulk. These losses can be due to tunneling and Ohmic resistance at the interface, as well as Ohmic resistance and other electrical losses within the bulk of the semiconductor. It is defined as

\begin{equation}
    \eta_{V_f} = \frac{E_{h\nu}}{V_f}
\end{equation}

where $E_{h\nu}$ denotes the photon energy and $V_f$ is the diode forward voltage \cite{schubert2018light}\cite{tsao2010solid}.

\subsubsection{Light Extraction Efficiency $\eta_{LE}$}

Device light-extraction efficiency (LEE) describes the  losses due to absorption in the material after electron-hole recombination and the associated emission of a photon. It is defined as

\begin{equation}
    \eta_{LE}=\frac{\text{\# of photons out}}{\text{\# of photons created}} = \frac{P_{opt}}{P_{int}}
\end{equation}

where $P_{opt}$ is the optical power of the device and $P_{int}$ is the internal power of the device\cite{schubert2018light}\cite{tsao2010solid}.

\subsubsection{Internal Quantum Efficiency $\eta_{IQ}$}

Internal quantum efficiency (IQE) describes non-radiative recombinatory processes in the semiconductor bulk. It is related to the external quantum efficiency through
\begin{equation}
\label{eqn:ieq-eqe}
    \eta_{EQ} = \eta_{IQ} \times \eta_{LE}
\end{equation}

and defined as

\begin{equation}
    \eta_{IQ} = \frac{\# \text{of photons created}}{\# \text{of electron-hole pairs in}}
\end{equation}

and depends on the current density of the device\cite{schubert2018light}\cite{tsao2010solid}.

\subsubsection{Droop $\eta_{droop}$}

Efficiency droop describes the decrease of device internal quantum efficiency at high current densities, which is caused by a number of different physical effects\cite{David2020}. This is often treated separately from internal quantum efficiency in the literature due to its importance in high-power devices. It is defined as

\begin{equation}
    \eta_{droop} = 1 - \frac{\eta_{IQE}}{\eta_{IQE}(A \rightarrow 0)}
\end{equation}

where $\eta_{IQE}(A \rightarrow 0)$ denotes the internal quantum efficiency at low current densities. In practice, droop is often given as the percentage difference between the ideal luminous intensity curve $\phi_{ideal}$ and the real luminous intensity curve $\phi$ at a set diode test current $A_{test}$ as

\begin{equation}
\label{eqn:droop}
    D = \frac{\phi_{ideal}(A_{test})-\phi(A_{test})}{\phi_{ideal}(A_{test})/100}
\end{equation}

Therefore, according to this definition, a lack of droop corresponds to a droop efficiency of $\eta_{droop} = 100\%$\cite{schubert2018light}\cite{tsao2010solid}.

\subsubsection{Conversion Efficiency $\eta_{C}$}

Light conversion efficiency (LCE) describes losses in the conversion process of blue light that is the basis for all phosphor-converted white LEDs. These losses are the sum of the Stokes loss and well as scattering/absorption losses. It is defined as

\begin{equation}
    \eta_{C} = \frac{E_{\textcolor{blue}{B}}}{\sum_{i=\textcolor{red}{R},\textcolor{orange}{O},\textcolor{yellow}{Y},\textcolor{teal}{G}} E_i}
\end{equation}

where $E_i$ denotes the total energy of light at the color corresponding to the down-converted photon wavelength \cite{schubert2018light}\cite{tsao2010solid}. Since every individual phosphor component in the device has its own associated conversion losses, the denominator sums over all components, i.e. \textcolor{red}{R}ed, \textcolor{orange}{O}range, \textcolor{yellow}{Y}ellow, \textcolor{teal}{G}reen.

\subsubsection{Spectral Efficiency $\eta_{S}$}

Spectral efficiency (SE) describes losses in the conversion process due to the wavelength-dependent efficiency of the human eyes. Photons converted into the infrared or ultraviolet are lost to illumination purposes. It is defined as

\begin{equation}
    \eta_{S} = \frac{K}{K_{max}(CRI,CCT)}
\end{equation}

where $K$ is the luminous efficacy of radiation of the light source, which can be computed from the device spectrum and the luminosity function, which describes the sensitivity of the human eye. $K_{max}$ is the maximum luminous efficacy of radiation of a perfect light source with the same color rendering performance and correlated color temperature as the light source in question\cite{schubert2018light}\cite{tsao2010solid}.

\newpage
\section{Methodology}
\label{sec:costmodel}

\subsection{Manufacturing Cost Model}

\subsubsection{Structure of the Model}

@@

\subsubsection{Manufacturing Process Steps by Chip Architecture}

@@

\subsubsection{Computation of Manufacturing Cost}
\label{subsec:cost_computation}

@@

\subsubsection{Computation of Yielded Cost}

@@

\newpage
\subsubsection{Example of Input Data Considered: Sapphire Wafers}

@@

\newpage
\subsubsection{Preliminary Sensitivity Analysis}

@@

\newpage
\subsubsection{Comparison with Reported Industry Data and DOE Projections}

@@

\newpage
\subsection{Computation of Contribution of Individual Variables}
\label{sec:contribution_variables}

@@

\section{LED Innovations}

\subsection{Non-Phosphor-Related}

@@

\subsection{Phosphor-Related}
\label{sec:innovation_phosphor}

@@

\subsubsection{History of Identified Phosphor-Related LED Innovations}

\paragraph{YAG and YGAG Phosphors}

@@

\paragraph{258, SLA and SALON phosphors}

@@

\paragraph{PFS Phosphor}

@@

\paragraph{Quantum Dots for Light Down-Conversion}

@@

\paragraph{Spectral Data for Identified LED Phosphor Innovations}

@@

\section{List of Interviews}

@@

\clearpage
\section{Results}

\subsection{Historical Progress in Sub-Efficiencies}

The historical progress in the different sub-efficiencies contributing to overall efficiency (lamp efficiency) is shown in \cref{fgr:subeff}. Context for figure panels is provided below:

\textbf{Panel A1:} Device forward voltage at a test current of I=350mA. The physical limit for a blue light wavelength of 450nm without electric pumping is shown for reference. Data points for devices released in 2020 by various manufacturers are shown in an inset plot. \

\textbf{Panel A2:} Efficiency droop at the test current of I=350mA. \

\textbf{Panel B1:} Internal quantum efficiency, for different chip architectures, by type of measurement used. Note that the artifact in the DOE Average around 2013 is due to a change in definition for internal quantum efficiency laid out in \cite{doe_ssl_multiyear_2013}. Measurement methods: PL - Photoluminescence\cite{Shim_2018}, EL - Electroluminescence\cite{Getty_2009};  ABC Model\cite{Karpov_2014}, Ideality Factor\cite{Masui_2010}. “Not reported” denotes data points where the IQE measurement technique was not reported. Whiskers indicate reported ranges, where applicable. \

\textbf{Panel B2:} Light extraction efficiency for different chip architectures, by type of specific technology used to improve light-extraction efficiency. The source of data points is shown for reference: simulation - ray-tracing computer simulation; experimental (absolute) - light-extraction efficiency given directly; experimenal (relative) - relative improvement over baseline chip architecture given. Abbreviations: TFFC - Thin-Film Flip-Chip; FC - Flip-Chip; CSP - Chip-Scale Package; ITO - Indium Tin Oxide; ZnO - Zinc Oxide; PSS - Patterned Sapphire Substrate; electr. - electrode; text. - textured.

\begin{figure}[h!]
 \centering
 \includegraphics[width=15.5cm]{2_SSL_EST/article/figures/subefficiencies_progress.pdf}
 \caption{Historical improvements in light-emitting diode technology. Panels A1-A2 show they key metrics used to compute forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show the sub-efficiencies internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. Source: own synthesis of data from the full list of sources provided in \cref{sec:sources}. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports \cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report \cite{beale_leds_2015}.}
 \label{fgr:subeff}
\end{figure}

\subsection{Manufacturing Cost Model}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=15.5cm]{2_SSL_EST/article/figures/costmodel_results_years.pdf}
 \caption{Manufacturing cost structure modelled for a single low-to-mid power GaN, classic chip, phosphor-converted white LED package, assuming an ideal factory with state-of-the-art equipment at a U.S. location. Consecutive steps shown on the x-axis. Panels A-C: Waterfall diagrams of LED manufacturing cost split by manufacturing process steps for years 2003, 2012 and 2020. Process steps on the horizontal axis are sequenced from left to right in the same order as in the modelled LED manufacturing process. Panel F: Cumulative manufacturing yield after each process step for years 2003, 2012 and 2020. For a visual representation of the overview of the manufacturing process, compare the diagrams in Section 2 of the Supplementary Information. Abbreviations: Litho – Lithographic Process, Insp. – Inspection, Depn. – Deposition, CMP - Chemical-Mechanical Planarization}
 \label{fgr:costmodel}
\end{figure*}
